---
title: "STAI Stability Analysis"
author: "Franziska Werner"
date: "29 10 2021"
output: word_document
papaja::apa6_docx: default
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: no
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: UKE 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r, load libraries and data files, message=FALSE}

library("papaja")
r_refs("r-references.bib")
library(car)
library(ggplot2)
library(knitr)
library(kableExtra)
library(psych)
#library(DescTools)
library(data.table)
library(plyr)
library(dplyr)
library(tidyverse)
library(jmv)
library(sjPlot)
library(ggpubr)
library(ggExtra)
library(lcsm)
library(zoo)
library(grid)
library(cowplot)
library(gridExtra)
library(apaTables)
library(forcats)
library(ez)
library(stringr)
library(PupillometryR)
library(lattice)
library(rmarkdown)
library(Rmisc)
library(gghalves)
library(devtools)
library(gghighlight)
library(lme4)
library(sjstats)
library(emmeans)
library(performance)
library(randomForest)
library(see)
library(here)

# load data files
load(here::here("data/StaiRetest.Rdata"))
load(here::here("data/B07_STAI_T.Rdata"))
load(here::here("data/B07_STAI_T_items.Rdata"))
load(here::here("data/B07_STAI_S.Rdata"))
load(here::here("data/B07_time_gaps.Rdata"))

```

```{r, complete dataset first measurement, warning=FALSE, message=FALSE, results='hide'}
# first data sample is a subset of a core data set with 690 participants
id.core <- matdata$staiRetest[[1]][[1]][,1]
items.core <- matdata$staiRetest[[1]][[1]][,2:21]
items.fMRT <- matdata$staiRetest[[2]][[1]][,2:21]
items.v1 <- matdata$staiRetest[[3]][[1]][,2:21]

tb.items.core <- as_tibble(items.core)

# generate dataframe for all 690 participants (first measuring point)
data.MP1 <- tibble(
  ID = id.core,
  tb.items.core
)

# rename columns to MP1_Item1...20
names(data.MP1)[2:ncol(data.MP1)] <- paste0(rep("MP1_Item", each = 20), 1:20)
```

```{r, generate dataset first sample, results='hide'}
# Pick people from 2nd measurement time point from the first (core) data set
fmrt.choice <- matdata$staiRetest[[2]][[2]][,2]
data.MP1[fmrt.choice,]

v1.choice <- matdata$staiRetest[[3]][[2]][,2]
data.MP1[v1.choice,]

# Connect fmrt und v1 to complete dataset (1st measurement point)
data.MP1.komplett <- data.MP1 %>%
  {rbind(.[fmrt.choice,], .[v1.choice,])} %>%
  {.[-c(31),]} # 31st person being excluded

# data (items) 2nd measuring point 
items.M2 <- as_tibble(rbind(items.fMRT, items.v1))
# rename columns
names(items.M2)[1:ncol(items.M2)] <- paste0(rep("MP2_Item", each = 20), 1:20)
data.MP2 <- items.M2[-c(31),]

data.M1_M2 <- as_tibble(cbind(data.MP1.komplett, data.MP2))

# recoding specific items 
data.R <- data.M1_M2 %>%
    mutate(across(c(MP1_Item1, MP1_Item6, MP1_Item7, MP1_Item10, MP1_Item13, MP1_Item16, MP1_Item19, MP2_Item1, MP2_Item6, MP2_Item7, MP2_Item10, MP2_Item13, MP2_Item16, MP2_Item19),
        ~recode(., `1` = 4, `2` = 3, `3` = 2, `4` = 1)
    ))

# data = complete dataframe for first sample: 
# OS_M1: Overall Score for first measurement, OS_M2: Overall Score for 2nd measurement
# MP1_Item1-20: Rating for specific item for measuring point 1
# MP2_Item1-20: Rating for specific item for measuring point 2
data <- tibble(
  ID = 1:nrow(data.R),
  ID.correspond = data.R$ID,
  OS_M1 = rowSums(data.R[, 2:21]),
  OS_M2 = rowSums(data.R[, 22:41]),
  data.R[2:ncol(data.R)]
)

```

```{r, generate dataset second sample, results='hide'}
tb.data <- as_tibble(B07_STAI_T)

# check number of NAs
sum(is.na(tb.data$STAIT_T3))

# exclude NAs
tb.data <- tb.data %>% filter(!is.na(STAIT_Z02), !is.na(STAIT_T3), !is.na(STAIT_T4), !is.na(STAIT_T5))

# renaming 
tb.data <- tb.data %>% rename(T1 = STAIT_Z02, T2 = STAIT_T3, T3 = STAIT_T4, T4 = STAIT_T5, T5 = STAIT_T5.2)
```

```{r, load STAI-S data and remove NAs from dataset}
# load STAI-S data (corresponding items are already inverted)
tb.STAI.S <- as_tibble(B07_STAI_S)

tb.STAI.S.w <- tb.STAI.S %>% select(id, timepoint, STAI_S_Sum) %>% pivot_wider(names_from = timepoint, values_from = STAI_S_Sum)

## remove NAs
tb.STAI.S.w.rmNA <- tb.STAI.S.w[complete.cases(tb.STAI.S.w), ]

```

```{r, ICC calculation for first sample}
# ICC Calculation with psych (detach("package:DescTools", unload=TRUE)) to have the confidence intervals calculated as well
icc <- c()
pVal.T1 <- c()
lowBound <- c()
upperBound <- c()

for (i in 1:20){
  dataframe <- data.frame(data[,i+4], data[,i+24])
  icc2 <- round(ICC(dataframe, missing = FALSE, lmer = FALSE)[[1]][2,2:8], 3)
  icc[i] <- icc2[[1]]
  pVal.T1[i] <- icc2[[5]]
  lowBound[i] <- icc2[[6]]
  upperBound[i] <- icc2[[7]]
}

dataframe.ICC <- data.frame(Item = 1:20, ICC = icc)

## ICC overall score
# generate dataframe for ICC calculation
df.tScore.icc <- data.frame(data$OS_M1, data$OS_M2)
# calculate ICC
icc.ts <- round(ICC(df.tScore.icc, missing = FALSE, lmer = FALSE)[[1]][2,2], 3) # ICC2 = 0.426
```

```{r, Cronbachs alpha, echo=FALSE}
# first sample, first measuring point 
ia1 <- data %>%
  select(MP1_Item1:MP1_Item20) %>%
  alpha()

cronbachs.Alpha.1 <- ia1$total$raw_alpha
alpha.minus.item.1 <- ia1$alpha.drop$raw_alpha  # alpha, if deleted
trennschärfe.1 <- ia1$item.stats$r.drop
schwierigkeit.1 <- ia1$item.stats$mean/4

tb1.internal.consistency <- round(tibble(
  Item = 1:20,
  Discrimination = trennschärfe.1,
  Difficulty = schwierigkeit.1
), digits = 3)

```

```{r}
# first sample, second measuring point
ia <- data %>%
  select(MP2_Item1:MP2_Item20) %>% 
  alpha()

cronbachs.Alpha <- ia$total$raw_alpha
alpha.minus.item <- ia$alpha.drop$raw_alpha
trennschärfe <- ia$item.stats$r.drop
schwierigkeit <- ia$item.stats$mean/4

tb2.internal.consistency <- round(tibble(
  Discrimination = trennschärfe,
  Difficulty = schwierigkeit, 
  ICC = icc,
  lowerBound = lowBound,
  upperBound = upperBound
), digits = 3)

```

```{r, results='hide'}
# conversion to long data format (second sample)

data.long <- tb.data %>%
  pivot_longer(cols = c(T1, T2, T3, T4, T5), names_to = "Time", values_to = "Score")

```

# Methods
 <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
@spielberger2010

## Participants
- trial 1: 105 participants
- trial 2: 87 participants 

-> add: age, sex and education?

## Material
- description of context in which data was collected

## Data analysis
<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->

Analysis was conducted in R (R Core Team, 2017) and figures were produced using the package ggplot2 (Wickham, 2009).

# Results


```{r, Boxplot first sample, results='hide'}
tb.long <- data %>% 
  select(OS_M1:OS_M2) %>%
  pivot_longer(c(OS_M1, OS_M2))

mp <- rep(c(1.0,2.0), times = 105)

tb.boxPlot <- tibble(
  Total.Score = tb.long,
  MP = mp
)

tb.boxPlot$MP <- as.factor(tb.boxPlot$MP)

bp_trial1 <- 
  ggplot(tb.boxPlot, aes(x = MP, y = Total.Score$value)) +
  geom_boxplot(color = "coral2", outlier.colour = "red", outlier.shape = 8, outlier.size = 2, notch = T, show.legend = F, lwd = .3) + 
  ylim(20, 60) +
  labs(title = "",x = "Measurement", y = "Total Score") +
  theme_apa() + theme(axis.title = element_text(size = 8), axis.text = element_text(size = 8))
```

```{r, Boxplot second sample}
# select subset and transform to long 
tb.data.long <- tb.data %>% 
  select(T1, T4) %>%
  pivot_longer(c(T1, T4))

tb.boxPlot2 <- tibble(
  Score = tb.data.long$value,
  Measurement = tb.data.long$name
)

# renaming for consistency (plot)
tb.boxPlot2$Measurement[ tb.boxPlot2$Measurement == "T1" ] <- 1
tb.boxPlot2$Measurement[ tb.boxPlot2$Measurement == "T4" ] <- 4

bp_trial2 <- 
  ggplot(tb.boxPlot2, aes(x = Measurement, y = Score)) +
  geom_boxplot(color = '#91D1C2B2', outlier.colour = "red", outlier.shape = 8, outlier.size = 2, notch = T, show.legend = F, lwd = .3) +
  ylim(20, 60) +
  labs(title="",x="Measurement", y = "") +
  theme_apa() + theme(title = element_text(size = 10), axis.title.x = element_text(size = 8), axis.text = element_text(size = 8), axis.line.y = element_blank(), axis.ticks.y = element_blank(), axis.text.y = element_blank()) 
```

```{r, Violin-Box-Scatter-Spaghetti-Plot Trial 1, warning=FALSE}
tbl.T1.wide <- data %>% rename("1" = OS_M1, "2" = OS_M2)

tbl.T1.long <- tbl.T1.wide %>% select(ID, "1", "2") %>%
  pivot_longer(cols = c("1", "2"), names_to = "Time", values_to = "Score")

tbl.T1.long$Time <- as.numeric(as.character(tbl.T1.long$Time))
```

```{r}
df.summary.T1 <- summarySE(tbl.T1.long, measurevar = "Score", groupvars = c("Time"))
```

```{r, Violin-Box-Scatter-Spaghetti-Plot 1st sample, warning=FALSE}
vbssp.Plt.T1 <- 
  ggplot(tbl.T1.long, aes(x = as.numeric(Time), y = Score), fill = Time) + 
  geom_point(data = tbl.T1.long %>% filter(Time == 1), aes(x = 1), color = 'coral2', size = .8, 
              alpha = .6, position = position_jitter(w = 0.1, h = 0)) +
  geom_point(data = tbl.T1.long %>% filter(Time == 2), aes(x = 2), color = 'coral2', size = .8, 
              alpha = .6, position = position_jitter(w = 0.1, h = 0)) +
  geom_line(aes(x = Time, group = ID), color = 'lightgray', alpha = .3) +
  geom_half_boxplot(data = tbl.T1.long %>% filter(Time == 1), aes(x = 1, y = Score), position = position_nudge(x = -.28), side = "l", outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, notch = T, lwd = .3) +
  geom_half_boxplot(data = tbl.T1.long %>% filter(Time == 2), aes(x = 2, y = Score), position = position_nudge(x = .28), side = "l", outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, notch = T, lwd = .3) + 
  geom_half_violin(data = tbl.T1.long %>% filter(Time == 1), aes(x = 1, y = Score), colour = NA, position = position_nudge(x = -.5), side = "l", fill = 'coral2', trim = FALSE, alpha=.6) +
  geom_half_violin(data = tbl.T1.long %>% filter(Time == 2), aes(x = 2, y = Score), colour = NA, position = position_nudge(x = .4), side = "r", fill = 'coral2', trim = FALSE, alpha = .6) + 
  geom_point(data = df.summary.T1, aes(x = Time, y = Score), color = "blue", size = 1.5, alpha = 1) +
  geom_line(data = df.summary.T1, aes(x = Time, y = Score), color = 'blue', alpha = 1) +
  geom_errorbar(data = df.summary.T1, aes(x = as.numeric(Time), y = Score, group = Time, ymin = Score-se, ymax = Score+se), width = .05, color = "blue") +
  labs(title = "Sample 1 ", x = "Time (41 months)", y = "STAI-T sum score") +
  ylim(20, 60) +
  theme_apa() +
  scale_x_continuous(breaks=c(1, 2), labels=c("T1", "T2"), limits=c(0, 3)) +
  theme(title = element_text(size = 10), axis.title = element_text(size = 9), axis.text = element_text(size = 9))
  
```

```{r, prepare data for violin-box-scatter-spaghettiplot for second sample}
# caution: T4 is now renamed to "2" due to later plotting reasons...
tbl.T2.wide <- tb.data %>% rename("1" = T1, "2" = T4)

tbl.T2.long <- tbl.T2.wide %>% select(id, "1", "2") %>%
  pivot_longer(cols = c("1", "2"), names_to = "Time", values_to = "Score")

tbl.T2.long$Time <- as.numeric(as.character(tbl.T2.long$Time))
```

```{r}
df.summary.T2 <- summarySE(tbl.T2.long, measurevar = "Score", groupvars = c("Time"))
```

```{r, Violin-Box-Scatter-Spaghetti-Plot 2nd sample, warning=FALSE}
vbssp.Plt.T2 <- ggplot(tbl.T2.long, aes(x = as.numeric(Time), y = Score), fill = Time) + 
  geom_point(data = tbl.T2.long %>% filter(Time == 1), aes(x = 1), color = '#91D1C2B2', size = .8, 
              alpha = .6, position = position_jitter(w = 0.1, h = 0)) +
  geom_point(data = tbl.T2.long %>% filter(Time == 2), aes(x = 2), color = '#91D1C2B2', size = .8, 
              alpha = .6, position = position_jitter(w = 0.1, h = 0)) +
  geom_line(aes(x = Time, group = id), color = 'lightgray', alpha = .3) +
  geom_half_boxplot(data = tbl.T2.long %>% filter(Time == 1), aes(x = 1, y = Score), position = position_nudge(x = -.28), side = "l", outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, notch = T, lwd = .3) +
  geom_half_boxplot(data = tbl.T2.long %>% filter(Time == 2), aes(x = 2, y = Score), position = position_nudge(x = .28), side = "l", outlier.shape = NA, center = TRUE, errorbar.draw = FALSE, width = .2, notch = T, lwd = .3) + 
  geom_half_violin(data = tbl.T2.long %>% filter(Time == 1), aes(x = 1, y = Score), colour = NA, position = position_nudge(x = -.5), side = "l", fill = '#91D1C2B2', trim = FALSE) +
  geom_half_violin(data = tbl.T2.long %>% filter(Time == 2), aes(x = 2, y = Score), colour = NA, position = position_nudge(x = .4), side = "r", fill = '#91D1C2B2', trim = FALSE) + 
  geom_point(data = df.summary.T2, aes(x = Time, y = Score), color = "blue", size = 1.5, alpha = .6) +
  geom_line(data = df.summary.T2, aes(x = Time, y = Score), color = 'blue', alpha = .8) +
  geom_errorbar(data = df.summary.T2, aes(x = as.numeric(Time), y = Score, group = Time, ymin = Score-se, ymax = Score+se), width = .05, color = "blue") +
  labs(title = "Sample 2 ", x = "Time (32 months)") +
  ylim(20, 60) +
  theme_apa() +
  scale_x_continuous(breaks=c(1, 2), labels=c("T1", "T4"), limits=c(0, 3)) +
  theme(title = element_text(size = 10), axis.title.x = element_text(size = 9), axis.title.y = element_blank(), axis.text = element_text(size = 9), axis.text.y = element_blank(), axis.line.y = element_blank(), axis.ticks.y = element_blank())

```

```{r, fig.width = 12, fig.height = 5, fig.cap = "Illustration of STAI-T scores for (A) Study 1 in red and (B) Study 2 in green which depicts the sample average (blue) at both time-points (A: 41 month apart; B: 32 month apart), individual data points (A: red dots, B: green dots)  as well as box and density plots at both time points. Note that the STAI-T scores of the same individual at both measurement time points are connected through a grey line.", warning=FALSE}
plot_grid(vbssp.Plt.T1, NULL, vbssp.Plt.T2, labels = c('A', '', 'B'), label_x = c(.02, 0, -.08), nrow = 1, rel_widths = c(1, .08, 1))
```

```{r}
tbl.ItemAnalysis <- cbind(tb1.internal.consistency, tb2.internal.consistency)
```

```{r, message=FALSE, results='hide'}
# transform columns of dataframe tbl.ItemAnalysis so that leading zeros are omitted
omitZeros <- c()
i = 1
numformat <- function(val) { sub("^(-?)0.", "\\1.", sprintf("%.2f", val)) }

for (k in 2:8){
  for (val in tbl.ItemAnalysis[,k]) {
    print(val)
    omitZeros[i] <- numformat(val)
    i <- i + 1
  }
  tbl.ItemAnalysis[,k] <- omitZeros
  omitZeros <- c()
  i = 1
  }
```


```{r, add significance level 1st sample, results='hide'}
# prepare data for table (first sample)
icc[1] <- ".26**"
icc[2] <- ".43**"
icc[3] <- ".26**"
icc[4] <- ".14"
icc[5] <- ".40**"
icc[6] <- ".04"
icc[7] <- ".32**"
icc[8] <- ".29**"
icc[9] <- ".36**"
icc[10] <- ".28**"
icc[11] <- ".41**"
icc[12] <- ".52**"
icc[13] <- ".35**"
icc[14] <- ".32**"
icc[15] <- ".35**"
icc[16] <- ".29**"
icc[17] <- ".29**"
icc[18] <- ".23**"
icc[19] <- ".39**"
icc[20] <- ".39**"

tbl.ItemAnalysis$ICC <- icc
```

```{r, results="asis"}
# generates table for first sample
apa_table(tbl.ItemAnalysis,
          align = c("c", rep("c", 8)),
          caption = "Item Analysis and item stability.",
          note =  "Total number of participants was $N = 105$. Cronbach's $\\alpha$ resulted in .87 and .89 for the first and the second measurement. \\\\ $^a$ The definition choice was absolute agreement. \\\\ *$p$ < .05   **$p$ < .01",   
          escape = FALSE,
          format.args = list(digits = c(0, 2, 2, 2, 2, 2, 2, 2)),
          #format = "latex",
          col.names = c("Item", "Discrimination", "Difficulty", "Discrimination", "Difficulty", "ICC(2,1)$^a$", "LL", "UL"),
          col_spanners = list(`Measurement 1` = c(2, 3), `Measurement 2` = c(4, 5), `95\\,\\% CI` = c(7, 8)))
          
```

```{r, generate dataset 2nd sample single item responses, results='hide'}

tb.allData <- as_tibble(B07_STAI_T_items)

## check number of NA's
sum(is.na(tb.allData))

## rename timepoints ##
# R base
# tb.allData$timepoint[ tb.allData$timepoint == "Z02" ] <- "T1"

tb.allData <- tb.allData %>% mutate( timepoint = ifelse(timepoint == "Z02","T1", timepoint ) )
tb.allData <- tb.allData %>% mutate( timepoint = ifelse(timepoint == "T3","T2", timepoint ) )
tb.allData <- tb.allData %>% mutate( timepoint = ifelse(timepoint == "T4","T3", timepoint ) )
tb.allData <- tb.allData %>% mutate( timepoint = ifelse(timepoint == "T5","T4", timepoint ) )
tb.allData <- tb.allData %>% mutate( timepoint = ifelse(timepoint == "T5.2","T5", timepoint ) )

## recode items
tb.allData.Re <- tb.allData %>%
    mutate(across(c(STAI1, STAI6, STAI7, STAI10, STAI13, STAI16, STAI19),
        ~recode(., `1` = 4, `2` = 3, `3` = 2, `4` = 1)
    ))

```

```{r, ICC calculation single items 2nd sample, message=FALSE}

## to consider just the 87 participants
sample2nd.T1 <- tb.allData.Re %>% subset(timepoint == "T1")
sample2nd.T1.red <- inner_join(sample2nd.T1, tb.data %>% select(id))

row.sums.T1 <- rowSums(sample2nd.T1.red[,3:22]) 

sample2nd.T4 <- tb.allData.Re %>% subset(timepoint == "T4")
sample2nd.T4.red <- inner_join(sample2nd.T4, tb.data %>% select(id))

row.sums.T4 <- rowSums(sample2nd.T4.red[,3:22]) # just to test whether I get same results as Maren

sample2nd.T1.T4 <- rbind(sample2nd.T1.red, sample2nd.T4.red)

icc.items.T2 <- c()
pV.T2 <- c()
lowBound.T2 <- c()
upperBound.T2 <- c()
tb.Items <- tb.allData.Re %>% select(starts_with("S"))
stai <- c(names(tb.Items))
k <- 1

for (item in stai) {
  
    tb.Cor <- sample2nd.T1.T4[, c("id", "timepoint", item)] %>%
    pivot_wider(names_from = timepoint, values_from = item) %>%
    select(T1, T4)
    
    icc.items.T2[k] <- round(ICC(tb.Cor, missing = FALSE, lmer = FALSE)[[1]][2,2], 2)
    pV.T2[k]        <- ICC(tb.Cor, missing = FALSE, lmer = FALSE)[[1]][2,6]
    lowBound.T2[k]  <- round(ICC(tb.Cor, missing = FALSE, lmer = FALSE)[[1]][2,7], 2)
    upperBound.T2[k]<- round(ICC(tb.Cor, missing = FALSE, lmer = FALSE)[[1]][2,8], 2)
    k <- k + 1
}

```

```{r, internal consistency 2nd sample 1st measurement, message=FALSE}
# 2nd sample, first measurement
ia.T2.M1 <- sample2nd.T1.T4 %>%
  subset(timepoint == "T1") %>%
  select(STAI1:STAI20) %>%
  alpha()

cronbachs.Alpha.T2.M1 <- ia.T2.M1$total$raw_alpha
alpha.minus.item.T2.M1 <- ia.T2.M1$alpha.drop$raw_alpha  # alpha, if deleted
discrimination.T2.M1 <- ia.T2.M1$item.stats$r.drop
difficulty.T2.M1 <- ia.T2.M1$item.stats$mean/4

tb1.internal.consistency.T2 <- round(tibble(
  Item = 1:20,
  Discrimination = discrimination.T2.M1,
  Difficulty = difficulty.T2.M1
), digits = 2)

```

```{r, internal consistency 2nd sample 2nd measurement, message=FALSE}
# 2nd sample, second measurement
ia.T2.M2 <- sample2nd.T1.T4 %>%
  subset(timepoint == "T4") %>%
  select(STAI1:STAI20) %>%
  alpha()

cronbachs.Alpha.T2.M2 <- ia.T2.M2$total$raw_alpha
alpha.minus.item.T2.M2 <- ia.T2.M2$alpha.drop$raw_alpha
discrimination.T2.M2 <- ia.T2.M2$item.stats$r.drop
difficulty.T2.M2 <- ia.T2.M2$item.stats$mean/4

tb2.internal.consistency.T2 <- round(tibble(
  Discrimination = discrimination.T2.M2,
  Difficulty = difficulty.T2.M2, 
  ICC = icc.items.T2,
  lowerBound = lowBound.T2,
  upperBound = upperBound.T2
), digits = 2)

```

```{r, message=FALSE}
tbl.ItemAnalysis.T2 <- cbind(tb1.internal.consistency.T2, tb2.internal.consistency.T2)
```

```{r, results='hide'}
# transform columns of dataframe tbl.ItemAnalysis so that leading zeros are omitted
omitZeros <- c()
i = 1
numformat <- function(val) { sub("^(-?)0.", "\\1.", sprintf("%.2f", val)) }

for (k in 2:8){
  for (val in tbl.ItemAnalysis.T2[,k]) {
    print(val)
    omitZeros[i] <- numformat(val)
    i <- i + 1
  }
  tbl.ItemAnalysis.T2[,k] <- omitZeros
  omitZeros <- c()
  i = 1
  }
```

```{r, significance level to ICC column, echo=FALSE}
# prepare data 2nd sample for table
icc.items.T2[1] <- ".48**" 
icc.items.T2[2] <- ".37**" 
icc.items.T2[3] <- ".30**"  
icc.items.T2[4] <- ".10"   
icc.items.T2[5] <- ".45**"  
icc.items.T2[6] <- ".22*"  
icc.items.T2[7] <- ".38**"  
icc.items.T2[8] <- ".30**"  
icc.items.T2[9] <- ".35**" 
icc.items.T2[10] <- ".49**" 
icc.items.T2[11] <- ".52**" 
icc.items.T2[12] <- ".47**" 
icc.items.T2[13] <- ".48**" 
icc.items.T2[14] <- ".13" 
icc.items.T2[15] <- ".18*" 
icc.items.T2[16] <- ".37**" 
icc.items.T2[17] <- ".37**" 
icc.items.T2[18] <- ".44**" 
icc.items.T2[19] <- ".49**" 
icc.items.T2[20] <- ".23**" 

tbl.ItemAnalysis.T2$ICC <- icc.items.T2

```

```{r, results="asis", message=FALSE}
# table 2nd sample
apa_table(tbl.ItemAnalysis.T2,
          align = c("c", rep("c", 8)),
          caption = "Item Analysis and item stability.",
          note =  "Total number of participants was $N = 87$. Cronbach's $\\alpha$ resulted in .87 and .92 for the first and the fourth measurement. \\\\ $^a$ The definition choice was absolute agreement. \\\\ *$p$ < .05   **$p$ < .01",   
          escape = FALSE,
          format.args = list(digits = c(0, 2, 2, 2, 2, 2, 2, 2)),
          col.names = c("Item", "Discrimination", "Difficulty", "Discrimination", "Difficulty", "ICC(2,1)$^a$", "LL", "UL"),
          col_spanners = list(`Measurement 1` = c(2, 3), `Measurement 2` = c(4, 5), `95\\,\\% CI` = c(7, 8)))
```

```{r, correlation plot 1st sample, message = F, warning = F}

ICC.T1 <- ICC(data %>% select(OS_M1, OS_M2), missing = F, lmer = F)

ICC2.T1 <- round(ICC.T1[[1]][2,2], 3)
p.T1    <- ICC.T1[[1]][2,6]
lowB.T1 <- round(ICC.T1[[1]][2,7], 3)
upB.T1  <- round(ICC.T1[[1]][2,8], 3)

plt.CorTS.T1 <-
  ggplot(data, aes(x = OS_M1, y = OS_M2)) + 
  geom_point(size = .3, color = "coral2") + 
  geom_smooth(aes(fill = "coral2"), method ="lm", se = TRUE, color = "coral2", size = .8, show.legend = F) +
  annotate("text", x = 32, y = 62, label = "ICC(2,1): abs. agreement = .43", size = 3) +
  ylim(20, 62) +
  labs(title = "1st sample: ICC Total Score", x = "First measurement", y = "Measurement after approx. 41 months") +
  theme_apa() + theme(title = element_text(size = 10), axis.title = element_text(size = 8), axis.text = element_text(size = 8))
```

```{r, correlation plot 2nd sample}

ICC.T2 <- ICC(tb.data %>% select(T1, T4), missing = F, lmer = F)

ICC2.T2 <- round(ICC.T2[[1]][2,2], 3)
p.T2    <- ICC.T2[[1]][2,6]
lowB.T2 <- round(ICC.T2[[1]][2,7], 3)
upB.T2  <- round(ICC.T2[[1]][2,8], 3)

plt.CorTS.T2 <- 
  ggplot(tb.data, aes(x = T1, y = T4)) + 
  geom_point(size = .3, col = '#91D1C2B2', alpha = 1) + 
  geom_smooth(method="lm", se = TRUE, fill = '#91D1C2B2', col = '#91D1C2B2', size = .8, show.legend = F) +
  annotate("text", x = 33, y = 62, label = "ICC(2,1): abs. agreement = .49", size = 3) +
  labs(title = "2nd sample: ICC Total Score", x = "First measurement", y = "Measurement after approx. 32 months") +
  ylim(20, 62) +
  theme_apa() + theme(title = element_text(size = 10), axis.title = element_text(size = 8), axis.text = element_text(size = 8))
```

```{r, fig.cap = "Retest Reliability. Shown are correlations in achieved total score between two measurements for sample one (A) and sample two (B). Single points represent individual participants.", fig.width=8, fig.height=4}

plot_grid(plt.CorTS.T1, plt.CorTS.T2, labels = c('A', 'B'))

```

```{r, results='hide', fig.cap = "Correlations in total score for the STAI-S. Number of participants was 72.", fig.width=6, fig.height=4, warning=FALSE}
## STAI-S Total Score Correlation
ICC.STAI.S <- ICC(tb.STAI.S.w.rmNA %>% select(T0, T5), missing = F, lmer = F)

ICC2.STAI.S <- round(ICC.STAI.S[[1]][2,2], 3)
p.STAI.S    <- ICC.STAI.S[[1]][2,6]
lowB.STAI.S <- round(ICC.STAI.S[[1]][2,7], 3)
upB.STAI.S  <- round(ICC.STAI.S[[1]][2,8], 3)

plt.CorTS.STAI.S <- 
  ggplot(tb.STAI.S.w.rmNA, aes(x = T0, y = T5)) + 
  geom_point(size = .3, col = 'black', alpha = 1) + 
  geom_smooth(method="lm", se = TRUE, fill = 'lightgray', col = 'black', size = .8, show.legend = F) +
  annotate("text", x = 28, y = 50, label = "ICC(2,1): abs. agreement = .275, p = 6.4e-04", size = 3) +
  labs(title = "STAI-S: ICC Total Score", x = "First measurement", y = "Fourth measurement") +
  theme_apa() + theme(title = element_text(size = 10), axis.title = element_text(size = 8), axis.text = element_text(size = 8))

```

```{r, generate dataset for first ANOVA (factors: time and item)}
tb.ANOVA <- tibble(
  score = c(rep(data$OS_M1, 20),rep(data$OS_M2, 20)),
  timepoint = factor(c(rep(1, 2100),rep(2,2100))),
  item = factor(c(rep(1, 105), rep(2, 105), rep(3, 105), rep(4, 105), rep(5, 105), rep(6, 105), rep(7, 105), rep(8, 105), rep(9, 105), rep(10, 105), rep(11, 105), rep(12, 105), rep(13, 105), rep(14, 105), rep(15, 105), rep(16, 105), rep(17, 105), rep(18, 105), rep(19, 105), rep(20, 105), rep(1, 105), rep(2, 105), rep(3, 105), rep(4, 105), rep(5, 105), rep(6, 105), rep(7, 105), rep(8, 105), rep(9, 105), rep(10, 105), rep(11, 105), rep(12, 105), rep(13, 105), rep(14, 105), rep(15, 105), rep(16, 105), rep(17, 105), rep(18, 105), rep(19, 105), rep(20, 105)))
)
```

```{r, results='hide'}
result.ANOVA <- aov(score ~ timepoint * item, data = tb.ANOVA)
summary(result.ANOVA)
anova.out <- apa_print(result.ANOVA)
```

The twofactorial ANOVA showed a temporal effect on the total score (`r anova.out$full_result$timepoint`), whereas the individual items had no effect (`r anova.out$full_result$item`).  

```{r, mixed model first sample}

id.vec <- rep(1:105, 40)
tb.ANOVA.ID <- tb.ANOVA %>% add_column(id = id.vec)

# dV: score, fixed effects: timepoint and item, does association exists after controlling for variation in test person: random effect: (1 | id)
fitS1 <- lmer(score ~ timepoint * item + (1 | id), data = tb.ANOVA.ID)
print(summary(fitS1))

# random effects part tells me how much variance I find among levels of my grouping factors, plus residual variance
#Random effects:
# Groups   Name        Variance Std.Dev.
# id       (Intercept) 40.45    6.360   
# Residual             16.85    4.105   
#Number of obs: 4200, groups:  id, 105

# dividing id variance by total variance (40.45/(40.45 + 16.85)) = 70% -> id explains a lot of variation: 
# differences between test persons explain approx. 70% of the variance thats left over AFTER!! the variance explained by our fixed effects!!!

### car als extension für lmer 
#m1 <- lmer(score ~ timepoint * item + (1 | id), data = tb.ANOVA.ID , 
#contrasts=list(timepoint = contr.sum, item = contr.sum )) #,type = 3)
#dist <- check_distribution(m1)
#plot(dist)
#anova(m1,type = '3')
#effectsize::omega_squared(m1)
#lsmeans(m1, pairwise ~ timepoint * item, adjust = "tukey")

```

```{r, generate dataset for ANOVA 2nd sample: Effect of Lockdown}
data.after <- tb.data %>% 
  subset(., tb.data$participated_lockdown == "after")
data.before <- tb.data[grep("before",tb.data$participated_lockdown),]
data.within <- tb.data %>% subset(participated_lockdown == "within")

# combine "within" datasets
d.within1 <- data.within[,c(1,4,7)]
d.within2 <- tb.data[grep("within/after",tb.data$participated_lockdown),][,c(1,4,7)]
d.within <- arrange(rbind(d.within2, d.within1), id)

tb.ANOVA2 <- tibble(
  score = c(data.before$T1, d.within$T1, data.after$T1, data.before$T4, d.within$T4, data.after$T4), 
  timePoint = factor(c(rep(1, 87), rep(4, 87))), 
  partLockdown = rep(c(rep("before", 46), rep("within", 19), rep("after", 22)), 2)
)
```

```{r, ANOVA, results='hide'}
result.ANOVA2 <- aov(score ~ timePoint * partLockdown, data = tb.ANOVA2)
summary(result.ANOVA2)
anova2.out <- apa_print(result.ANOVA2)
```

The twofactorial ANOVA showed that it made no difference in terms of total score whether participants completed the STAI-T before, during, or after lockdown (`r anova2.out$full_result$partLockdown`). Therefore, further analysis is done without differentiating between the three subgroups. Instead, we consider the complete data set.  

```{r, mixed model 2nd sample: Did the lockdown influence the achieved score?}

id.vec.2 <- rep(1:87, 2)
tb.ANOVA2.ID <- tb.ANOVA2 %>% add_column(id = id.vec.2)

m2 <- lmer(score ~ timePoint * partLockdown + (1 | id), data = tb.ANOVA2.ID , 
contrasts=list(timepoint = contr.sum, partLockdown = contr.sum )) 
dist <- check_distribution(m2)
plot(dist)
anova(m2, type = '3')
effectsize::omega_squared(m2)
lsmeans(m2, pairwise ~ timePoint * partLockdown, adjust = "tukey")

```

```{r, ICC Calculation}
# this functions computes respective ICC values and fills in the result table
calcICC <- function(tableICC, data_ICC, k){

    results_ICC <- ICC(data_ICC, missing = FALSE, lmer = FALSE)
    
    ## fill results table 
    tableICC$ICC2_agreement[k] <- round(results_ICC[[1]][2,2], 3)
    tableICC$ICC2_lowerCI[k] <- round(results_ICC[[1]][2,7], 3)
    tableICC$ICC2_upperCI[k] <- round(results_ICC[[1]][2,8], 3)
    tableICC$ICC2_p_value[k] <- round(results_ICC[[1]][2,6], 6)
    
    return(tableICC)
}
```

```{r}
### Create empty matrix for results
ICC_multipleTP <- data.frame(matrix(data = NA, nrow = 6, ncol = 5, byrow = TRUE, dimnames = NULL))
colnames(ICC_multipleTP) <- c("combinations", "ICC2_agreement","ICC2_lowerCI","ICC2_upperCI","ICC2_p_value")
```

```{r}
# for complete dataset (without discrimination between participated lockdown)
k <- 1
times1 <- c("T1", "T2", "T3")
times2 <- c("T1", "T2", "T3", "T4")

  for (j in 1:length(times1)) { 
    for (l in (j+1):length(times2)) {

    data_ICC <- data.frame(data.long$Score[which(data.long$Time == times1[j])], data.long$Score[which(data.long$Time == times2[l])])
    
    ICC_multipleTP <- calcICC(ICC_multipleTP, data_ICC, k)
    k <- k + 1
    
    }
  }

combinations <- c(rep(c("1,2", "1,3", "1,4", "2,3", "2,4", "3,4")))

ICC_multipleTP$combinations <- combinations
```

```{r,fig.cap = "ICC for different samples and different time intervals with 95 % confidence intervals. The gray entry indicates the results for the STAI-S, the red one represents the results for the first sample (STAI-T) and the turquoise one represents the results for the second sample (STAI-T) with its different time intervals. Missing data in brackets indicate that there were only two mesaurements in total.", fig.width=6, fig.height=5}
# generates multiple ICC plot
# we only consider complete data sets (without T5)
elapsed.Time <- c("32 (1,4)", "26 (1,3)", "20 (1,2)", "12 (2,4)", "6 (3,4)", "5 (2,3)")

grob <- grobTree(textGrob("poor", x=0.06,  y=0.98, hjust=0,
  gp=gpar(col="gray50", fontsize=8, fontface="italic")))

grob1 <- grobTree(textGrob("moderate", x=0.43,  y=0.98, hjust=0, 
  gp=gpar(col="gray50", fontsize=8, fontface="italic")))

grob2 <- grobTree(textGrob("good", x=0.73,  y=0.98, hjust=0, 
  gp=gpar(col="gray50", fontsize=8, fontface="italic")))

grob3 <- grobTree(textGrob("excellent", x=0.84,  y=.98, hjust=0, 
  gp=gpar(col="gray50", fontsize=8, fontface="italic")))

## new order for later displaying 
tb.ICC.ordered <- ICC_multipleTP[c(3,2,1,5,6,4),]

ICC2_table <- tb.ICC.ordered %>%
  add_column(pastTime = elapsed.Time, 
             .after = "combinations") %>% select(pastTime, ICC2_agreement, ICC2_lowerCI, ICC2_upperCI)

# add STAI-S data to dataframe
ICC2_tableSTAI <- ICC2_table %>% add_row(pastTime = "31 (1,4)", ICC2_agreement = ICC2.STAI.S, ICC2_lowerCI = lowB.STAI.S, ICC2_upperCI = upB.STAI.S, .after = 1)

# add STAI-T first sample to dataframe 
ICC2_tableAll <- ICC2_tableSTAI %>% add_row(pastTime = "41", ICC2_agreement = ICC2.T1, ICC2_lowerCI = lowB.T1, ICC2_upperCI = upB.T1, .before = 1)

# add groups to dataframe for later coloring
df.ICCs <- ICC2_tableAll %>% add_column(group = c("sample1", "sample2", "STAI-S", "sample2", "sample2", "sample2", "sample2", "sample2"), 
                                        .before = "pastTime")

# to make R respect the order in the dataframe
df.ICCs$pastTime <- factor(df.ICCs$pastTime, levels = df.ICCs$pastTime[order(df.ICCs$ICC2_agreement)])

# assign color to different groups
color = rep(NA, length=length(df.ICCs$group))
color[which(df.ICCs$group=="sample1")] = "coral2"
color[which(df.ICCs$group=="sample2")] = "#91D1C2B2"
color[which(df.ICCs$group=="STAI-S")] = "gray50"

  ggplot(df.ICCs, aes(x=ICC2_agreement,y=pastTime, color=color)) +
  geom_point(show.legend = F, color = color) + 
  geom_errorbar(aes(xmin=ICC2_lowerCI, xmax=ICC2_upperCI), width=.5, color = color, show.legend = F) +

  geom_vline(xintercept = 0.0, color = "grey50", size = 1, linetype = "dotted") + 
  geom_vline(xintercept = 0.5, color = "grey50", size = 1, linetype = "dotted") +
  geom_vline(xintercept = 0.75, color = "grey50", size = 1, linetype = "dotted") +
  geom_vline(xintercept = 0.9, color = "grey50", size = 1, linetype = "dotted") +
  annotation_custom(grob) +
  annotation_custom(grob1) +
  annotation_custom(grob2) +
  annotation_custom(grob3) +
    scale_y_discrete(limits = c("5 (2,3)", "6 (3,4)", "12 (2,4)", "20 (1,2)", "26 (1,3)", "31 (1,4)", "32 (1,4)", "41")) +
  labs(title = "", x = "ICC(2,1): absolute agreement", y = "Average elapsed time [months] with considered measuring points") +
  theme_apa() + theme(axis.title = element_text(size = 8), axis.text = element_text(size = 8))

```























